---
title: "Ad influence survey analysis"
author: "Srijan Nayak"
date: "12/12/2020"
output: pdf_document
urlcolor: blue
papersize: a4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      cache = TRUE)
```

# Objective

The objective of this [survey](https://forms.gle/791dhPnxgLVPzgyG9) was to
understand the influence ads have on people's purchase decisions and on what
aspect of ads do people focus on the most.

# Required libraries

The following `R` libraries are used for this analysis.

```{r}
library(dplyr) # for wrangling data
library(ggplot2) # for making plots
library(readr) # for reading data
```

# Reading the data

```{r}
ad_influence_data <- read_csv("ad-influence.csv") %>%
  select(!Timestamp) # ignore timestamp column
n <- dim(ad_influence_data)[1]
questions <- names(ad_influence_data)
```

At the time of writing `r n` responses have been collected.

# Simplifying responses

Some options for the questions were made a little bit more descriptive to make
it easier for the user to select the option they felt appropriate. But these
long responses are not crucial for the analysis, therefore these are simplified
to shorter responses.

```{r}
simplify_responses <- function(response) {
  if (response == "Social media sites (like Facebook, Instagram)") {
    response = "Social media sites"
  } else if (response == "Video streaming sites (like YouTube, Vimeo)") {
    response = "Video streaming sites"
  } else if (response == "General web (like Google Ads on various sites)") {
    response = "General web"
  } else if (grepl("Recommendations? from friends or family", response)) {
    response = "Recommendations"
  }
  response
}

for (i in 1:length(ad_influence_data))  {
  ad_influence_data[questions[i]] <- ad_influence_data[questions[i]] %>%
    pull() %>%
    sapply(simplify_responses, USE.NAMES = FALSE)
}
```

# Platform with the most ads

The first questions asked in the survey was as follows: \
**_`r questions[1]`_**

Let's look at the top 5 responses for this question.

```{r fig.height=3}
ad_influence_data %>%
  rename(most_ads_platform = questions[1]) %>%
  group_by(most_ads_platform) %>%
  summarise(proportion = n() / n) %>%
  ungroup() %>%
  arrange(desc(proportion)) %>%
  head(5) %>%
  ggplot(aes(x = reorder(most_ads_platform, proportion), y = proportion)) +
  geom_col(width = 0.7) +
  labs(x = NULL, title = "Platforms with highest ad encounter rates") +
  coord_flip() +
  theme_classic()
```

It can be observed that nowadays people encounter ads the most on video
streaming sites and social media sites rather than traditional platforms like
television and radio channels. This suggests that people might be gradually
moving away from traditional media like televisions and FM radio.

# Platforms that draw the most attention towards ads

The second question in the survey was as follows: \
**_`r questions[2]`_**

The purpose of this question was to see which platform makes the user focus
on ads the longest. Let's look at the top 5 responses for this question.

```{r fig.height=3}
ad_influence_data %>%
  rename(attention_grabbing_platform = questions[2]) %>%
  group_by(attention_grabbing_platform) %>%
  summarise(proportion = n() / n) %>%
  ungroup() %>%
  arrange(desc(proportion)) %>%
  head(5) %>%
  ggplot(aes(x = reorder(attention_grabbing_platform, proportion),
             y = proportion)) +
  geom_col(width = 0.7) +
  labs(x = NULL, y = "Proportion",
       title = "Platforms where users pay attention to ads the most") +
  coord_flip() +
  theme_classic()
```

While the ranking of the categories is the same as what was observed in the
previous question responses, the difference in the proportions of the 
top 3 responses is much smaller. Thus video streaming sites, social media sites,
as well as television channels are good for advertising as the users are more
likely to pay attention to the ads showed on these platforms.

# What matters the most?

The third question in the survey was as follows: \
**_`r questions[3]`_**

The purpose of this question was to see if ads are even necessary any more,
and if they are, then what aspect is of the ad matters the most.

Let's look at the responses for this question.

```{r fig.height=3}
ad_influence_data %>%
  rename(important_aspect = questions[3]) %>%
  group_by(important_aspect) %>%
  summarise(proportion = n() / n) %>%
  ungroup() %>%
  arrange(desc(proportion)) %>%
  ggplot(aes(x = reorder(important_aspect, proportion),
             y = proportion)) +
  geom_col(width = 0.7) +
  labs(x = NULL, y = "Proportion",
       title = "Most important factor in making a purchase decision") +
  coord_flip() +
  theme_classic()
```

Looking at the plot, it appears that people prioritize recommendations and their
own research over ads. Brand ambassadors seem to be least popular factor for
a user making a purchase decision.

# How are new products discovered?

The third question in the survey was as follows: \
**_`r questions[4]`_**

Let's look at how the responses for this question turned out.

```{r fig.height=2}
ad_influence_data %>%
  rename(discover_source = questions[4]) %>%
  group_by(discover_source) %>%
  summarise(proportion = n() / n) %>%
  ungroup() %>%
  arrange(desc(proportion)) %>%
  ggplot(aes(x = reorder(discover_source, proportion),
             y = proportion)) +
  geom_col(width = 0.7) +
  labs(x = NULL, y = "Proportion",
       title = "How are new porducts/services discovered?") +
  coord_flip() +
  theme_classic()
```

It appears that people mostly discover products/services they are interested in
through recommendations or searching and browsing on their own instead of ads.
