---
title: "Ad influence survey analysis"
author: "Srijan Nayak"
date: "12/12/2020"
output: pdf_document
urlcolor: blue
papersize: a4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Objective

The objective of this [survey](https://forms.gle/791dhPnxgLVPzgyG9) was to
understand the influence ads have on people's purchase decisions and on what
aspect of ads do people focus on the most.

# Required libraries

The following `R` libraries are used for this analysis.

```{r}
library(dplyr) # for wrangling data
library(ggplot2) # for making plots
library(readr) # for reading data
```

# Reading the data

```{r}
ad_influence_data <- read_csv("ad-influence.csv") %>%
  select(!Timestamp) # ignore timestamp column
n <- dim(ad_influence_data)[1]
questions <- names(ad_influence_data)
```

At the time of writing `r n` responses have been collected.

# Simplifying responses

Some options for the first 2 questions were made a little bit more descriptive
to make it easier for the user to select the option they felt appropriate. But
these long responses are not crucial for the analysis, therefore these are
simplified into shorter responses.

```{r}
simplifyResponses <- function(response) {
  if (response == "Social media sites (like Facebook, Instagram)") {
    response = "Social media sites"
  } else if (response == "Video streaming sites (like YouTube, Vimeo)") {
    response = "Video streaming sites"
  } else if (response == "General web (like Google Ads on various sites)") {
    response = "General web"
  }
  return(response)
}

ad_influence_data[questions[1]] <- sapply(pull(ad_influence_data[questions[1]]),
                                          simplifyResponses, USE.NAMES = FALSE)
ad_influence_data[questions[2]] <- sapply(pull(ad_influence_data[questions[2]]),
                                          simplifyResponses, USE.NAMES = FALSE)
```
